Quantifying passive kelp rafting phenomena and connectivity of nearshore communities around the coast of South Africa:Development of a 3D spectral numerical coastal-flow model

### 2019-11-28 literature_review.Rmd
lit review updated

---
title: "Quantifying passive kelp rafting phenomena and connectivity of nearshore communities around the coast of South Africa:Development of a 3D spectral numerical coastal-flow model"
author: 
- RM Coppin^[University of the Western Cape]
- C Rautenbach^[South African Weather Service]
- AJ Smit^[University of the Western Cape]
output:
  pdf_document:
    toc: no
  html_document:
    df_print: paged
    toc: yes
    theme:  simplex
  word_document:
    fig_caption: yes
    reference_docx: chapter_1_template.docx
    toc: yes
bibliography: references.bib
---

@@

## Introduction
# Hydrodynamic modelling
## Introduction

Wave exposure may be modeled through various methods which range from simple cartographic to more advanced numerical wave models. Traditional ecological measures of wave exposure usually incorporates integrative measures of hydrodynamic conditions at a particular site. Cartographic models can be qualitative or quantitative and were designed for the need of wave exposure measures to explain ecological distributions. A simple set of calculations on coastline and wind data, and relatively small input data sets are required. These are regarded as "fetch-based models", which measure the length of open water associated with a site along a straight line. The output of such an approach is a simplified estimate of the potential wave energy for a specific set of sites. Advances in cartographical methods using fetch-based models has allowed for wave exposure measurements for larger areas, and has been suggested as a method for predicting macroalgal community structure (???). An example of such a model is the “BioEx model” which was developed by Baardseth and others (1970) to estimate wave exposure over large regions. BioEx requires frequency, strength and direction of winds, weighted by degree of exposure within various directions. BioEx is calculated as the sum of the index developed at different spatial scales (local, fjord and open), and has been used in mapping marine coastal biodiversity (???). Lindegarth and Gamfeldt (2005) criticised this approach, arguing that the choice of wave exposure method can influence ecological inference. The authors also highlighted the need for objective, reproducible and quantitative studies comparing exposure indices (Lindegarth and Gamfeldt 2005). Other authors, such as Hill et al. (2010), have argued that these simple measures can be improved upon by including bathymetry data which allows the incorporation of diffraction into the calculation. Diffraction is topographically induced variations in wave direction. A model incorporating this complexity was developed by Isæus (2004), and is known as the “simplified wave model” (SWM). The model uses measurements of wind strength, fetch and empirically derived algorithms to mimic diffraction.

Advances in numerical modelling have been founded on physical wave theory which describes how a wave "behaves". This approach is based on a theoretical perspective rather than the need to answer ecological questions. Besides diffraction, numerical models incorporate more complexity by including wind forcing, wave-to-wave interactions and loss of energy due to friction and wave breaking. Numerical models have a variety of applications and are often incorporated within hydrodynamic general circulation models and are used operationally for forecasting the sea state (Group 1988; Booij, Ris, and Holthuijsen 1999; Smith, Sherlock, and Resio 2001). The downside of advanced numerical models is that are computationally intensive which creates limitations for large scale simulations. Therefore, their application along long stretches of variable coastline, inshore environments and ocean-wide simulations are limited due to the poor spatial coverage. However, numerical models can be designed for local or site specific coverage, provided the correct data is available. **...More to be added...**

>>>>

---
title: "Quantifying passive kelp rafting phenomena and connectivity of nearshore communities around the coast of South Africa:Development of a 3D spectral numerical coastal-flow model"
author: 
- RM Coppin^[University of the Western Cape]
- C Rautenbach^[South African Weather Service]
- AJ Smit^[University of the Western Cape]
output:
  html_document:
    df_print: paged
    toc: yes
    theme:  simplex
  word_document:
    fig_caption: yes
    reference_docx: chapter_1_template.docx
    toc: yes
bibliography: references.bib
---

@@

# Hydrodynamic modelling
## Introduction

Wave exposure may be modeled through various methods which range from simple cartographic to more advanced numerical wave models. Traditional ecological measures of wave exposure usually incorporates integrative measures of hydrodynamic conditions at a particular site. Cartographic models can be qualitative or quantitative and were designed for the need of wave exposure measures to explain ecological distributions. A simple set of calculations on coastline and wind data, and relatively small input data sets are required. These are regarded as "fetch-based models", which measure the length of open water associated with a site along a straight line. The output of such an approach is a simplified estimate of the potential wave energy for a specific set of sites. Advances in cartographical methods using fetch-based models has allowed for wave exposure measurements of larger areas, and has been suggested as a method for predicting macroalgal community structure (???). An example of such a model is the “BioEx model” which was developed by Baardseth and others (1970) to estimate wave exposure over large regions. BioEx requires frequency, strength and direction of winds, weighted by degree of exposure within various directions. BioEx is calculated as the sum of the index developed at different spatial scales (local, fjord and open), and has been used in mapping marine coastal biodiversity (???). Lindegarth and Gamfeldt (2005) criticised this approach, arguing that the choice of wave exposure method can influence ecological inference. The authors also highlighted the need for objective, reproducible and quantitative studies comparing exposure indices (Lindegarth and Gamfeldt 2005). Other authors, such as Hill et al. (2010), have argued that these simple measures can be improved upon by including bathymetry data which allows the incorporation of diffraction into the calculation. Diffraction is topographically induced variations in wave direction. A model incorporating this complexity was developed by Isæus (2004), and is known as the “simplified wave model” (SWM). The model uses measurements of wind strength, fetch and empirically derived algorithms to mimic diffraction. Advances in numerical modelling have been founded on physical wave theory which describes how a wave "behaves". This approach is based on a theoretical perspective rather than the need to answer ecological questions. Besides diffraction, numerical models incorporate more complexity by including wind forcing, wave-to-wave interactions and loss of energy due to friction and wave breaking. Numerical models have a variety of applications and are often incorporated within hydrodynamic general circulation models and are used operationally for forecasting the sea state (Group 1988; Booij, Ris, and Holthuijsen 1999; Smith, Sherlock, and Resio 2001). The downside of advanced numerical models is that are computationally intensive which creates limitations for large scale simulations. Therefore, their application along long stretches of variable coastline, inshore environments and ocean-wide simulations are limited due to the poor spatial coverage. However, numerical models can be designed for local or site specific coverage, provided the correct data is available. Most models have been designed with a specific goal or objective in mind, and so no single model exists that is appropriate for all situations or applications. Although models are "situation specific", completed models can set the foundation for future models with regards to resolving important oceanographic processes, identifying gaps, and identifying important tracer coefficients [@North2009]. Modelling of nearshore ecosystems has been reviewed by @Jones2002, and offers the reader a general overview of the discipline (**I must try and get this paper**).

## Practicalities of model design

Hydrodynamic models use the same primitive equations for ocean models, which calculate velocities, turbulence, temperature, and salinity, and hence, density [@North2009]. The equations used in hydrodynamic modelling need to be 'discretised', which means the equations are formulated to be evaluated at discrete temporal and spatial points. Additionally, hydrodynamic models use similar approaches to grid and coordinate structures (see ocean modelling section).
